The complex network of trophic interactions in a subAntarctic oceanic Marine Protected Area

### 2024-03-27 MS_ToSubmit/RevisedVersion_2.Rmd
Revised version 2

Maintaining ocean ecosystem services relies on a healthy ocean, and biological species play a key role in this process. It is essential to identify the main species involved in sustaining this health for effective management and conservation purposes. In this regard, @Bergagna2020 have identified the principal ecosystem services that are provided by species inhabiting the MPA N-BB; for instance, nursery and feeding areas for commercially important fishes (e.g. Patagonian grenadier *Macruronus magellanicus*, Patagonian toothfish *Dissostichus eleginoides*), water purification by filter feeders (e.g. sponges), among others. These species are part of a complex system in terms of biodiversity. There is robust knowledge on the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022]. Overall, 811 benthic and plankton species have been identified in the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. Recently, the structure of the southwestern South Atlantic Ocean has been hypothesised to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure, meaning a shorter average food chain and a greater trophic overlap and redundancy, respect to other sub-Antarctic areas such as the continental shelf off Tierra del Fuego. The high abundance of a limited number of prey species at the intermediate trophic level of the food web which leads to a specific feeding reliance among high-trophic level predators. This phenomenon shortens food-web length while simultaneously amplifying trophic overlap and redundancy within the overall food web. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* has been considered the most plausible 'wasp-waist' species for the MPA N-BB [@Riccialdelli2020], since they are species with high regional abundances [@Madirolas2000; @Arkhipkin2013], they occupy mid-trophic levels [@Riccialdelli2010; @Riccialdelli2020], many predators feed on them [@Riccialdelli2020], and their population dynamics appear to depend on the environment (e.g. climatic variations) [@Diez2018]. These recent efforts have increased our understanding of the biological communities and their complexity, but an understanding of species interactions’ complexity and structure is lacking. Such understanding will provide a more complete picture of the ecosystem, generate a baseline to comprehend climate change effects [@Franco2020] and serve as an input to decision makers on conservation actions [@AdministraciondeParquesNacionales2022].

This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The assembly of these interactions in a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem. In recent years, topological network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats to the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and exploitation of gas and oil) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

In this work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a topological network approach to a highly resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in the network. For the former, we considered all species and interactions in the food web, and calculated the number of species (S), number of interactions (L), link density (L/S), connectance (L/S^2), proportion of omnivorous, and assessed the small-world pattern. For the latter objective, we considered interactions and species related to a particular species and analyse the following properties: betweenness, closeness, trophic similarity, topological role and trophic level.

@@

Omnivory acts as a buffer to changes as the ecosystem presents alternative energy pathways in the face of perturbations, i.e., reducing the risk of cascading extinctions following the primary loss of species [@Borrvall2000]. This is supported by the fact that omnivorous adapt at a faster rate and to a wide range of environmental conditions due to its flexibility to feed on the most abundant prey [@Fagan1997]. Furthermore, omnivory can be analysed from the interaction point of view: theoretical studies have identified omnivorous interactions as a possible candidate for a keystone interaction, sensu @Kadoya2018, highlighting the importance of omnivory in stabilizing food web dynamics [@McCann1997; @Neutel2002]. The high proportion of omnivory in the food web of the MPA N-BB suggests that the network might be robust to variations in prey abundances, which could increase food web's persistence and stability [@Stouffer2010].

@@

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species would show similar trophic habits [@Cury2000]. Such trophic redundancy is expected to be greater for species in TLs above the waist that depend on a very few species at this intermediate level. However, our results show a negative relationship between trophic similarity (or redundancy) and trophic level, indicating functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web. This contradiction might be reflecting the potential discrepancy that could arise between qualitative (this study) and quantitative food web studies. Here it is important to note that our analyses considered all potential predator-prey interactions in the food web, assigning equal importance to each of them. Although this could be a disadvantage when inferring on energy flows, one of the main advantages of qualitative food web studies is that it considers the diversity of the communities inhabiting a particular study area, therefore providing a more representative picture of all components (species) acting on the ecosystem. In this sense, our study highlights the uniqueness in terms of the trophic role of high-TL predators. Not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp.*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].

>>>>

Maintaining ocean ecosystem services relies on a healthy ocean, and biological species play a key role in this process. It is essential to identify the main species involved in sustaining this health for effective management and conservation purposes. In this regard, @Bergagna2020 have identified the principal ecosystem services that are provided by species inhabiting the MPA N-BB; for instance, nursery and feeding areas for commercially important fishes (e.g. Patagonian grenadier *Macruronus magellanicus*, Patagonian toothfish *Dissostichus eleginoides*), water purification by filter feeders (e.g. sponges), among others. These species are part of a complex system in terms of biodiversity. There is robust knowledge on the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @AdministraciondeParquesNacionales2022]. Overall, 811 benthic and plankton species have been identified in the MPA N-BB ecosystem, where 349 were reported for the first time in the area in recent years [@AdministraciondeParquesNacionales2022]. Recently, the structure of the southwestern South Atlantic Ocean has been hypothesised to be under a ‘wasp-waist’ control, meaning that the structure and dynamics of the ecosystem are regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced ‘wasp-waist’ structure, meaning a shorter average food chain and a greater trophic overlap and redundancy in high-trophic levels compared to other sub-Antarctic areas, such as the continental shelf off Tierra del Fuego [@Riccialdelli2020]. The high abundance of a limited number of prey species at the intermediate trophic level of the food web leads to a specific feeding reliance among high-trophic level predators. This phenomenon shortens the length of the food web, as the diet of top predators is expected to be dominated by the mentioned intermediate-trophic level species, thus lowering their trophic position. Simultaneously, it amplifies trophic overlap and redundancy between such predators. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* have been considered the most plausible ‘wasp-waist’ species for the MPA N-BB [@Riccialdelli2020], since they are species with high regional abundances [@Madirolas2000; @Arkhipkin2013], they occupy mid-trophic levels [@Riccialdelli2010; @Riccialdelli2020], many predators feed on them [@Riccialdelli2020], and their population dynamics appear to depend on the environment (e.g. climatic variations) [@Diez2018]. These recent efforts have increased our understanding of the biological communities and their complexity, but an understanding of species interactions’ complexity and structure is lacking. Such understanding will provide a more complete picture of the ecosystem, generate a baseline to comprehend climate change effects [@Franco2020] and serve as an input to decision makers on conservation actions [@AdministraciondeParquesNacionales2022].

This aspect can be tackled by analysing one of the most frequent relationships between species: the predator-prey interaction [@Bascompte2009]. The assembly of these interactions in a particular region is referred to as a food web, representing the roadmap for matter and energy flow in an ecosystem. Here it is important to distinguish the inference power of topological food web models, which rely on presence/absence interactions among species, versus the so-called dynamic food web models that account for biomass of species, along with the magnitude and direction of energy flows among them [@Brose2009]. While the latter might better capture the ecosystem properties given that food webs are inherently dynamical systems, the former ones are the necessary first step, providing a map of biological and ecological-constrained possibilities. In recent years, topological network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among the anthropogenic threats reported for the MPA N-BB ecosystem, it is worth mentioning contaminants like mercury and microplastics [@Cossi2021; @DiMauro2022; @Fioramonti2022]; also fishing vessels are allowed to operate in the western section of the MPA N-BB (i.e. Marine National Reserve category), altering the stocks of commercially important fish species [@Martinez2021; @AdministraciondeParquesNacionales2022]. Moreover, there is a potential hazard related to the effects of offshore activities (exploration and exploitation of gas and oil) to the west of the MPA N-BB [@AdministraciondeParquesNacionales2022].

In this work, we present the first detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a topological network approach to a highly resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species’ role in the network. For the former, we considered all species and interactions in the food web, and calculated the number of species (S), number of interactions (L), link density (L/S), connectance (L/S2), proportion of omnivorous, and assessed the small-world pattern. For the latter objective, we considered interactions and species related to a particular species and analyse the following properties: betweenness, closeness, trophic similarity, topological role and trophic level.

@@

Omnivory acts as a buffer to changes as the ecosystem presents alternative energy pathways in the face of perturbations, i.e., reducing the risk of cascading extinctions following the primary loss of species [@Borrvall2000]. This is supported by the fact that omnivores adapt at a faster rate and to a wide range of environmental conditions due to its flexibility to feed on the most abundant prey [@Fagan1997]. Furthermore, omnivory can be analysed from the interaction point of view: theoretical studies have identified omnivorous interactions as a possible candidate for a keystone interaction, sensu @Kadoya2018, highlighting the importance of omnivory in stabilizing food web dynamics [@McCann1997; @Neutel2002]. The high proportion of omnivory in the food web of the MPA N-BB suggests that the network might be robust to variations in prey abundances, which could increase food web's persistence and stability [@Stouffer2010].

@@

Ecosystems with a pronounced 'wasp-waist' structure are suggested to present a high trophic redundancy, since many species feeding at intermediate trophic levels would show similar trophic habits [@Saporiti2015]. Such trophic redundancy is expected to be greater for species in TLs above the waist that depend on a very few species at this intermediate level. However, our results show a negative relationship between trophic similarity (or redundancy) and trophic level, indicating functional similarity at low and mid-TL species compared to higher TL species for the MPA N-BB food web. This contradiction might be reflecting the potential discrepancy that could arise between qualitative (this study) and quantitative food web studies. Here it is important to note that our analyses considered all potential predator-prey interactions in the food web, assigning equal importance to each of them. Although this could be a disadvantage when inferring on energy flows, one of the main advantages of qualitative food web studies is that it considers the diversity of the communities inhabiting a particular study area, therefore providing a more representative picture of all components (species) acting on the ecosystem. In this sense, our study highlights the uniqueness in terms of the trophic role of high-TL predators. Not only the expected pelagic animals such as marine mammals and seabirds arise as relevant, but also demersal vertebrate (chondrichthyans *Bathyraja macloviana* and *Squalus acanthias*) and benthic invertebrate species (echinoderms *Diplopteraster clarki* and *Pteraster sp.*) are noteworthy. The role that such species play in the MPA N-BB ecosystem is unique and perturbations on them might result in unprecedented changes at the trophic structure and functioning level. In this regard, we should mention the potential threat of the fisheries operating in the western section of the MPA N-BB, where this activity is allowed and mostly focuses on the Patagonian toothfish *Dissostichus eleginoides* and the southern blue whiting *Micromesistius australis* [@Martinez2015]. Although the fishing effort is concentrated outside the limits of the MPA N-BB, the impact on the MPA ecosystem should not be neglected [@Martinez2021].
